---
title: "scater"
author: "longlong"
date: "2022/2/9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load data, normalization and fliter data
```{r}
rm(list = ls()) 
Sys.setenv(R_MAX_NUM_DLLS=999)
## 首先载入文章的数据
load(file='../input.Rdata')
counts=a
library(stringr) 
meta=df

library(scater)
# 创建 scater 要求的对象
sce <- SingleCellExperiment(
  assays = list(counts = as.matrix(counts)), 
  colData = meta
)
## normalization
exprs(sce) <- log2(
  calculateCPM(sce ) + 1)
## 只有运行了下面的函数后才有各式各样的过滤指标
genes=rownames(rowData(sce))
genes[grepl('^MT-',genes)]
genes[grepl('^ERCC-',genes)]
## QC for colData-cells, add an information about ERCC
scecol <- perCellQCMetrics(sce, 
                         subsets = list(ERCC = grep('^ERCC',genes)))
## QC for rowData-genes, likely no useful
scerow=perFeatureQCMetrics(sce)
tmp1=data.frame(scerow)
## filter gene and sample by myself code
test1=assay(sce)
test2=apply(test1, 1, function(x) sum(x>0))

tmp=data.frame(scecol)

      
sce1=sce[test2>5,tmp$detected>2000]
```


# 基因表达，理论上应该是跟384孔板 这个变量无关
```{r}
plotExpression(sce1, rownames(sce1)[1:6],
               x = "plate", 
               exprs_values = "logcounts") 
```
# 展示高表达量基因, 绘图有点耗时
```{r}
plotHighestExprs(sce1, exprs_values = "counts")
```


# PCA
```{r}
# draw PCA
sce <- runPCA(sce1)
plotPCA(sce)
reducedDimNames(sce)
# colnames(as.data.frame(colData(sce)))
head(colData(sce))
## PCA分布图上面添加临床信息--------------
plotReducedDim(sce, dimred = "PCA", 
                shape_by= "plate", 
                colour_by= "g")
```

if(F){
## 考虑 ERCC 影响后继续PCA. Note: new method can not for col
sce@colData$plate
sce2 <- runPCA(sce1, 
                     subset_row = scerow$detected)
plotPCA(sce2)
## PCA分布图上面添加临床信息--------------
plotReducedDim(sce2, dimred = "PCA", 
               shape_by= "plate", 
               colour_by= "g")

}

# 运行 tSNE 降维算法
# !!!!!!!!! the bigger perplexity is, the more neighborhood number is. 
```{r}
set.seed(1000)
sce <- runTSNE(sce, perplexity=10)
plotTSNE(sce, 
         shape_by= "plate", 
         colour_by= "g")

```

# 对tSNE降维后结果进行不同的聚类
# !!!!!! Note: PCA, UMAP, tSNE are dimensional reduction methods. kmeans and hcluster are cluster methods
```{r}
colData(sce)$tSNE_kmeans <- as.character(kmeans(sce@int_colData@listData$reducedDims$TSNE,
                                                centers = 4)$clust)

hc=hclust(dist(sce@int_colData@listData$reducedDims$TSNE ))
plot(hc)
clus = cutree(hc, 4) 
colData(sce)$tSNE_hc <-  as.character(clus)
plotTSNE(sce,  colour_by = "tSNE_kmeans")
plotTSNE(sce,  colour_by = "tSNE_hc")
table(colData(sce)$tSNE_hc , colData(sce)$tSNE_kmeans)

## dimensional reduction by DiffusionMap
library("destiny")

sce1 <- runDiffusionMap(sce)

colData(sce1)$Diff_kmeans <- as.character(kmeans(sce1@int_colData@listData$reducedDims$DiffusionMap,
                                                centers = 4)$clust)
plotDiffusionMap(sce1,  
                 shape_by= "plate", 
                 colour_by= "Diff_kmeans")

## dimensional reduction by MDS
sce2=logNormCounts(sce)
sce1 <- runMDS(sce2)
plotMDS(sce1)
```


# SC3 package with scater
library(SC3) # BiocManager::install('SC3')
sce <- sc3_estimate_k(sce1)
## evaluate how many number of cluster
metadata(sce)$sc3$k_estimation
## preprocess before sc3 command
rowData(sce)$feature_symbol=rownames(rowData(sce))
# 耗费时间
kn=4
sc3_cluster="sc3_4_clusters"
# 非常耗时
sce <- sc3(sce, ks = kn, biology = TRUE)

sc3_plot_consensus(sce, k = kn, show_pdata = c(sc3_cluster))
sc3_plot_expression(sce, k = kn, show_pdata =  c(sc3_cluster))
sc3_plot_markers(sce, k = kn, show_pdata =  c(sc3_cluster))
plotPCA(sce, shape_by= "g" , colour_by =  sc3_cluster )
sc3_interactive(sce)





